**Title:** Analyzing long-term water quality of lakes in Rhode Island and the Northeastern United States with an anomaly approach

### 2020-09-30 manuscript/ecosphere_resubmit/manuscript.Rmd
few small changes - title was a bit off.

The analysis approach we used here, site-specific anomalies, is not a novel method and does have a long history in the analysis of trends in climate [@hansen2006global; @hansen2010global; @jones1996calculating; @jones1999surface]. However, using it to examine water quality trends is a new application of the technique, as we could find little evidence of using it specifically for water quality trends.  We built on these methods and adapted them for use with long-term water quality trends.  While other methods are valid and robust [e.g., @Oliver_2017], we chose median site-specific anomalies as they can provide readily interpretable results, especially for communicating to general audiences.  For instance, reporting the changes in anomalies allows us to look at changes in the original units.  With our analysis, the slope of the regression line for temperature suggests an average yearly increase of `r signif(ww_temp_rl$slope, 2)` °C and the slope of the regression line for chlorophyll *a* shows an average yearly increase of `r signif(ww_chla_rl$slope, 2)` µg/l.

>>>>

The analysis approach we used here, site-specific anomalies, is not a novel method and does have a long history in the analysis of trends in climate [@hansen2006global; @hansen2010global; @jones1996calculating; @jones1999surface]. However, using it to examine water quality trends is a little-used application of the technique, as we only found a single study using anomalies in lakes and that study's focus was only on lake temperature [@oreilly2015warming]. There is little evidence of using anomalies more broadly with water quality trends.  Thus, we built on these methods and adapted them for use with long-term water quality trends.  While other methods are valid and robust [e.g., @Oliver_2017], we chose median site-specific anomalies as they can provide readily interpretable results, especially for communicating to general audiences.  For instance, reporting the changes in anomalies allows us to look at changes in the original units.  With our analysis, the slope of the regression line for temperature suggests an average yearly increase of `r signif(ww_temp_rl$slope, 2)` °C and the slope of the regression line for chlorophyll *a* shows an average yearly increase of `r signif(ww_chla_rl$slope, 2)` µg/l.
